Edge sign prediction in general graphs and Correlation Clustering

### 2017-10-31 edge/signed.tex
pass over edge model and methods

\begin{prop}
  Let $\mathcal{D}_2 = \{w_1^\star, w_2^\star\}$ be an optimal solution of \autoref{p:edge_full} on the
  instance $\mathcal{I}$. Then $\mathcal{D}_2$ provides the following optimal solution to
  \autoref{p:edge_esp}: each every edge explained by $w_1^\star$ is predicted positive, and every
  edge explained by $w_2^\star$ is predicted negative.
\end{prop}
\begin{proof}
  Take any $w_1, w_2 \in \dsphere \times \dsphere$ such that $w_1 \neq w_2$. Assume without loss of
  generality that $w_1 = \argmax_{w\in \{w_1, w_2\}} {s_{\mathrm{inner}}}^T w$. This implies that:
  \begin{equation*}
    {s_{\mathrm{inner}}}^T w_1 = \frac{1}{d}\sum_{i=1}^d {w_1}_{;i} >
    \frac{1}{d}\sum_{i=1}^d {w_2}_{;i} = {s_{\mathrm{inner}}}^T w_2 \,.
  \end{equation*}
  In turn, we have that:
  \begin{equation*}
    {s_{\mathrm{outer}}}^T w_2 = -\frac{1}{d}\sum_{i=1}^d {w_2}_{;i} >
    -\frac{1}{d}\sum_{i=1}^d {w_1}_{;i} = {s_{\mathrm{outer}}}^T w_1 \,.
  \end{equation*}
  We can thus rewrite the maximization objective \eqref{eq:edge_full} of \autoref{p:edge_full} as:
  \begin{align}
    \label{eq:edge_special}
    \argmax_{w_1,w_2 \in \dsphere} & \quad \sum_{u,v \in E} \max_{w \in \{w_1, w_2\}} {s_{uv}}^T w = \\
    \argmax_{w_1,w_2 \in \dsphere} & \quad
    \sum_{\substack{u,v \in E\\ s_{uv} = s_{\mathrm{inner}}}} {s_{\mathrm{inner}}}^T w_1 +
    \sum_{\substack{u,v \in E\\ s_{uv} = s_{\mathrm{outer}}}} {s_{\mathrm{outer}}}^T w_2 = \notag \\
    \argmax_{w_1,w_2 \in \dsphere} & \quad m_i \frac{1}{d} \onev + m_o \frac{-1}{d} \onev \notag
  \end{align}
  One can check using Lagrange multipliers that for any $c \in \Rbb^n$, $\max_{w \in \dsphere} c^T w
  = \frac{c}{\|c\|}$. The solution of \eqref{eq:edge_special} is therefore:
  \begin{equation*}
    w_1^\star = \frac{\onev}{\sqrt{d}} \qquad \text{and} \qquad w_2^\star = \frac{-\onev}{\sqrt{d}}
  \end{equation*}
  As claimed, the edges associated with $w_1^\star$ (respectively $w_2^\star$) are the edges that
  are to be predicted positive (respectively negative) in \autoref{p:edge_esp}.
\end{proof}

\begin{aside}
  It is natural to ask how this result can be extended to more than $K=2$ groups. The most natural
  idea is to use one-hot encoding of the group membership as profiles, with additional negative
  values as in the example of \autoref{fig:edge_exe}. Namely, set the profiles of the \ith{} group
  as follow: all component are equal to $-b$ except for the \ith{} one, which is equal to $a$, where
  $a$ and $b$ are real constants such that $a^2 + (K-1)b^2=1$. However, in that case, we check
  numerically that $w_+ = \frac{\onev}{\sqrt{d}}$ and $w_- = \frac{-\onev}{\sqrt{d}}$ is the not the
  optimal solution, and that a solution achieving a larger objective function value does not cluster
  the edges between inner and outer.

>>>>

\begin{prop}
  Let $\mathcal{D}_2 = \{w_1^\star, w_2^\star\}$ be an optimal solution of \autoref{p:edge_full} on the
  instance $\mathcal{I}$. Then $\mathcal{D}_2$ provides the following optimal solution to
  \autoref{p:edge_esp}: each every edge explained by $w_1^\star$ is predicted positive, and every
  edge explained by $w_2^\star$ is predicted negative.
\end{prop}
\begin{proof}
  Take any $w_1, w_2 \in \dsphere \times \dsphere$ such that $w_1 \neq w_2$. Assume without loss of
  generality that $w_1 = \argmax_{w\in \{w_1, w_2\}} {s_{\mathrm{inner}}}^T w$. This implies that:
  \begin{equation*}
    {s_{\mathrm{inner}}}^T w_1 = \frac{1}{d}\sum_{i=1}^d {w_1}_{;i} >
    \frac{1}{d}\sum_{i=1}^d {w_2}_{;i} = {s_{\mathrm{inner}}}^T w_2 \,.
  \end{equation*}
  In turn, we have that:
  \begin{equation*}
    {s_{\mathrm{outer}}}^T w_2 = -\frac{1}{d}\sum_{i=1}^d {w_2}_{;i} >
    -\frac{1}{d}\sum_{i=1}^d {w_1}_{;i} = {s_{\mathrm{outer}}}^T w_1 \,.
  \end{equation*}
  We can thus rewrite the maximization objective \eqref{eq:edge_full} of \autoref{p:edge_full} as:
  \begin{align}
    \label{eq:edge_special}
    \argmax_{w_1,w_2 \in \dsphere} & \quad \sum_{u,v \in E} \max_{w \in \{w_1, w_2\}} {s_{uv}}^T w = \\
    \argmax_{w_1,w_2 \in \dsphere} & \quad
    \sum_{\substack{u,v \in E\\ s_{uv} = s_{\mathrm{inner}}}} {s_{\mathrm{inner}}}^T w_1 +
    \sum_{\substack{u,v \in E\\ s_{uv} = s_{\mathrm{outer}}}} {s_{\mathrm{outer}}}^T w_2 = \notag \\
    \argmax_{w_1,w_2 \in \dsphere} & \quad m_i {\frac{\onev}{d}}^T w_1  + m_o {\frac{-\onev}{d}} ^T w_2 \notag
  \end{align}
  One can check using Lagrange multipliers that for any $c \in \Rbb^n$, $\max_{w \in \dsphere} c^T w
  = \frac{c}{\|c\|}$. The solution of \eqref{eq:edge_special} is therefore:
  \begin{equation*}
    w_1^\star = \frac{\onev}{\sqrt{d}} \qquad \text{and} \qquad w_2^\star = \frac{-\onev}{\sqrt{d}}
  \end{equation*}
  As claimed, the edges associated with $w_1^\star$ (respectively $w_2^\star$) are the edges that
  are to be predicted positive (respectively negative) in \autoref{p:edge_esp}.
\end{proof}

\begin{aside}
  It is natural to ask how this result can be extended to more than $K=2$ groups. The most natural
  idea is to use one-hot encoding of the group membership as profiles, with additional negative
  values as in the example of \autoref{fig:edge_exe}. Namely, set the profiles of the \ith{} group
  as follow: all component are equal to $-b$ except for the \ith{} one, which is equal to $a$, where
  $a$ and $b$ are real constants such that $a^2 + (K-1)b^2=1$. However, in that case, we check
  numerically that $w_+ = \frac{\onev}{\sqrt{d}}$ and $w_- = \frac{-\onev}{\sqrt{d}}$ is the not the
  optimal solution, and that a solution achieving a larger objective function value does not cluster
  the edges between inner and outer groups.
